---
title: "Markdown_project_PAPE"
author: "nilsm"
date: "2025-11-04"
output: html_document
---

# Loading the data

To load the data correctly, the R-project file should be placed in dir containing:
1) a repertory "traffic_dataset" containing all the data parquet files ("results_2014.parquet")
2) the ref_ring.RData file

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Load packages
library(dplyr)
library(ggplot2)
library(sf) # library used for maps
library(arrow) # library used to read parquet files
library(lubridate) # library used to process dates and times
library(jsonlite)
library(osmdata)
library(leaflet)
library(mapedit)
library(stringr)
library(fixest)
library(rlang)


# This function loads the dataset between date_min and date_max (give them in format 'yyyy-mm-dd')
# If filter_na = TRUE, the function will not retrieve observations with no data either on flow or on occupancy rate
# If short = TRUE, the function will not retrieve the geometrical information (lighter for memory use so by default use this option!)

loading_function = function(date_min, date_max, filter_na = TRUE,short=TRUE) {
  
  if (short == TRUE) {
    filtered_cols = c("iu_ac","libelle", "t_1h", "q", "k")
  } else {
    filtered_cols = c("iu_ac", "libelle", "iu_nd_amont", "iu_nd_aval","t_1h", "q", "k", "etat_trafic", "etat_barre", "geometry")
  }
  
 res <- ds %>%
  select(all_of(filtered_cols)) %>%
  collect() %>%
  mutate(t_1h = ymd_hms(t_1h)) %>%
  filter(as.Date(t_1h) %within% interval(ymd(date_min), ymd(date_max)),
         month(t_1h) != 8)
  
  cat("Number of observations: ",nrow(res))
  
  return(res)
}



# Useful informations:
vehicle_length = 0.0045 # vehicle length mean is 4.5 meters

load('ref_ring.RData') #This RData file contains the geographical referential for the roads used in the analysis

gc()
  
  
```


# Computing arc roads
```{r}

geojson_to_sfg <- function(geojson) {
  geo <- fromJSON(geojson)
  coords <- geo$coordinates
  coords <- matrix(coords, ncol = 2)
  st_linestring(coords)  # retourne un objet sfg
}

convert_to_sf <- function(df, col) {
  df_name <- deparse(substitute(df))   # récupère le nom (ex: "VGP")
  
  geoms <- lapply(df[[col]], geojson_to_sfg)
  geoms <- st_sfc(geoms, crs = 4326)
  result <- st_sf(df, geometry = geoms)
  
  assign(paste0(df_name, "_sf"), result, envir = .GlobalEnv)
}

compute_arc_length = function(arc) {
  return(st_length(st_transform(arc, 32631))/1000)
}

ref_local_roads = rbind(Bd_St_Germain,Upper_bank,Lower_bank) %>% convert_to_sf(.,"geo_shape")
ref_ring_roads = rbind(PI_control_clean,PI_north_control_clean,PE_north_control_clean,PE_control_clean)

ref_local_roads$length = compute_arc_length(ref_local_roads)
ref_ring_roads$length = compute_arc_length(ref_ring_roads)


```


# Speed function

In the model, the speed is defined as $S(i,t) = \frac{Flow(i,t) * (L + Ki)}{Occupancy(i,t)}$ 
```{r}

compute_speed = function(df) {
  df$speed = df$q *(vehicle_length + as.numeric(df$length))/df$k
}

```




# Computing the model on ring road
```{r}

ds <- open_dataset('./data/ring.parquet')

categories_function_ring = function(df){
  Final = df %>% 
    mutate(treated = libelle %in% PE_control_clean$Libelle|libelle %in% PI_north_control_clean$Libelle) %>% # TRUE if arc is in the treated subset
    mutate(post = ymd(as.Date(t_1h)) > ymd("2016-09-01")) %>% # After ou before the VGP closure
    mutate(year_rel = floor(time_length(ymd(as.Date(t_1h)) - ymd("2016-08-31"), "year"))) %>% # year relative to VGP closure
    mutate(log_flow = log(q),log_occupancy = log(k)) %>% # Log of the flow and occupancy of cars
    mutate(date = as.Date(t_1h)) %>%
    mutate(wday = wday(ymd(date))) %>% # Day of the week (1 = Saturday, 2 = Monday, ..., 7 = Sunday)
    mutate(hours = hour(t_1h)) %>% # Hour of the day (0 to 23)
    mutate(day_hour = paste0(wday,'_',hours)) %>% # Wednesday at 8AM = "4_8" and Sunday at 7PM =  "7_19"
    mutate(north = libelle %in% PI_north_control_clean$Libelle|libelle %in% PE_north_control_clean$Libelle)
  
  Final$daytime = ifelse(Final$hours %in% c(8:20),TRUE,FALSE)
  Final$morning = ifelse(Final$hours %in% c(8:10),TRUE,FALSE)
  Final$evening = ifelse(Final$hours %in% c(18:20),TRUE,FALSE)
  
  return(Final)
}


date_min = "2013-09-01"
date_max = "2019-09-01"
df_ring = loading_function(date_min,date_max) %>% 
  filter(iu_ac %in% c(PE_north_control_clean$Identifiant.arc, PE_control_clean$Identifiant.arc,PI_control_clean$Identifiant.arc,PI_north_control_clean$Identifiant.arc))

Final = categories_function_ring(df_ring)

Final = Final %>% filter((is.na(q) & is.na(k))==F)

did_model <- feols(
  q ~ treated * post | libelle + t_1h,
  data = Final,
  cluster = ~libelle
)

summary(did_model)


# A function to compute the event model:
#   - Final is the data frame with the data filtered and with the categories
#   - subset is a CONDITION input ! e.g. "north == TRUE"
#   - variable to compute the model, e.g. "q", "k", "log_flow", "log_occupancy"...

compute_event_model = function(Final,subset,variable) {
  
  if (!is.null(subset)){
    df = Final %>% filter(!!parse_expr(subset))
  } else {df = Final}
  
  event_model <- feols(
    k ~ i(year_rel, treated, ref = "-1") | libelle + day_hour,
    data = df,
    cluster = ~libelle
  )
  
  summary(event_model)
  
  betas <- coef(event_model)
  ses <- se(event_model)
  
  results<- data.frame(
    coef_name = names(betas),
    beta = betas,
    se = ses,
    ci_low = betas - 1.96 * ses,
    ci_high = betas + 1.96 * ses
  )

  results = results %>% mutate(year_rel = sapply(coef_name,function(x) {return(unlist(str_split(x,":"))[3])}))
  results[6,] = c("year_rel::-1:treated",0,0,0,0,-1) # Add a line for the reference year (-1) with an estimate and se = 0
  results = results %>% select(-coef_name) %>% mutate(across(everything(),as.numeric))
  
  return(results)
}

results_all = compute_event_model(Final,subset = NULL,k)
results_south = compute_event_model(Final, subset = "north == FALSE",k)

results_south$subset = "South ring roads"
results_all$subset = "Full sample"


results = rbind(results_south,results_all)

ggplot(results, aes(x = year_rel, y = beta, color = subset)) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray") +
  geom_point(size = 2, position = position_dodge(width = 0.3)) +
  geom_errorbar(aes(ymin = ci_low, ymax = ci_high), width = 0,position = position_dodge(width = 0.3)) +
  geom_line(position = position_dodge(width = 0.3))+
  labs(
    x = "Year relative to VGP closure",
    y = "Estimates (β_k)",
    title = "Effect of VGP closure on car occupancy rate"
  ) +
  theme_minimal() + 
  scale_color_manual(values = c("Full sample" = "black", "South ring roads" = "darkgrey"))

```



# Computing the model on local road
```{r, include=FALSE}
ds <- open_dataset('./data/local.parquet')

categories_function_local = function(df){
  Final = df %>% 
    mutate(treated = libelle %in% Bd_St_Germain$Libelle|libelle %in% Upper_bank$Libelle) %>% # TRUE if arc is in the treated subset
    mutate(post = ymd(as.Date(t_1h)) > ymd("2016-09-01")) %>% # After ou before the VGP closure
    mutate(year_rel = floor(time_length(ymd(as.Date(t_1h)) - ymd("2016-08-31"), "year"))) %>% # year relative to VGP closure
    mutate(log_flow = log(q),log_occupancy = log(k)) %>% # Log of the flow and occupancy of cars
    mutate(date = as.Date(t_1h)) %>%
    mutate(wday = wday(ymd(date))) %>% # Day of the week (1 = Saturday, 2 = Monday, ..., 7 = Sunday)
    mutate(hours = hour(t_1h)) %>% # Hour of the day (0 to 23)
    mutate(day_hour = paste0(wday,'_',hours)) %>% # Wednesday at 8AM = "4_8" and Sunday at 7PM =  "7_19"
    mutate(road = case_when(libelle %in% Lower_bank$Libelle ~ "LB",
                            libelle %in% Upper_bank$Libelle ~ "UB",
                            libelle %in% Bd_St_Germain$Libelle ~ "BSG",
                            ))
  
  Final$daytime = ifelse(Final$hours %in% c(8:19),TRUE,FALSE)
  Final$morning = ifelse(Final$hours %in% c(8:9),TRUE,FALSE)
  Final$evening = ifelse(Final$hours %in% c(18:19),TRUE,FALSE)
  
  return(Final)
}

date_min = "2013-09-01"
date_max = "2019-09-01"
df_local = loading_function(date_min,date_max)

df_tot = rbind(df_ring,df_local)

Final = categories_function_local(df_local)

k_c = list()
for (i in unique(Final$iu_ac)){
  df = Final %>% filter(iu_ac == i, is.na(q) == F & is.na(k) == F)
  if(nrow(df) == 0){k_c[i]=NA
    }else{
    k_c[i] = -coef(lm(q ~ 0 + k + I(k^2), data = df))[1]/(2*coef(lm(q ~ 0 + k + I(k^2), data = df))[2])
    }
}

Final$kc = k_c[Final$iu_ac]
Final$congestion = ifelse(Final$k > Final$kc,1,0)

did_model <- feols(
  q ~ treated * post | libelle + t_1h,
  data = Final,
  cluster = ~libelle
)

summary(did_model)


# A function to compute the event model:
#   - Final is the data frame with the data filtered and with the categories
#   - subset is a CONDITION input ! e.g. "north == TRUE"
#   - variable to compute the model, e.g. "q", "k", "log_flow", "log_occupancy"...

compute_event_model = function(Final,subset,variable) {
  
  if (!is.null(subset)){
    df = Final %>% filter(!!parse_expr(subset))
  } else {df = Final}
  
  event_model <- feols(
    k ~ i(year_rel, treated, ref = "-1") | libelle + day_hour,
    data = df,
    cluster = ~libelle
  )
  
  summary(event_model)
  
  betas <- coef(event_model)
  ses <- se(event_model)
  
  results<- data.frame(
    coef_name = names(betas),
    beta = betas,
    se = ses,
    ci_low = betas - 1.96 * ses,
    ci_high = betas + 1.96 * ses
  )

  results = results %>% mutate(year_rel = sapply(coef_name,function(x) {return(unlist(str_split(x,":"))[3])}))
  results[6,] = c("year_rel::-1:treated",0,0,0,0,-1) # Add a line for the reference year (-1) with an estimate and se = 0
  results = results %>% select(-coef_name) %>% mutate(across(everything(),as.numeric))
  
  return(results)
}

results_all = compute_event_model(Final,subset = NULL,k)
results_south = compute_event_model(Final, subset = "north == FALSE",k)

results_south$subset = "South ring roads"
results_all$subset = "Full sample"


results = rbind(results_south,results_all)

ggplot(results, aes(x = year_rel, y = beta, color = subset)) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray") +
  geom_point(size = 2, position = position_dodge(width = 0.3)) +
  geom_errorbar(aes(ymin = ci_low, ymax = ci_high), width = 0,position = position_dodge(width = 0.3)) +
  geom_line(position = position_dodge(width = 0.3))+
  labs(
    x = "Year relative to VGP closure",
    y = "Estimates (β_k)",
    title = "Effect of VGP closure on car occupancy rate"
  ) +
  theme_minimal() + 
  scale_color_manual(values = c("Full sample" = "black", "South ring roads" = "darkgrey"))

```


```{r}


compute_model_plot = function(variable){
  
  fml <- as.formula(paste0(variable, " ~ i(year_rel, treated, ref = '-1') | libelle + day_hour"))
  
  event_model_daytime <- feols(
    fml,
    data = Final %>% filter(daytime == TRUE, road %in% c("BSG","UB","LB"), wday %in% c(2:6)),
    cluster = ~libelle
  )
  event_model_evening <- feols(
    fml,
    data = Final %>% filter(evening == TRUE, road %in% c("BSG","UB","LB"), wday %in% c(2:6)),
    cluster = ~libelle
  )
  event_model_morning <- feols(
    fml,
    data = Final %>% filter(morning == TRUE, road %in% c("BSG","UB","LB"), wday %in% c(2:6)),
    cluster = ~libelle
  )
  
  
  betas_morning <- coef(event_model_morning)
  ses_morning <- se(event_model_morning)
  
  results_morning <- data.frame(
    coef_name = names(betas_morning),
    beta = betas_morning,
    se = ses_morning,
    ci_low = betas_morning - 1.96 * ses_morning,
    ci_high = betas_morning + 1.96 * ses_morning
  )
  
  betas_evening <- coef(event_model_evening)
  ses_evening <- se(event_model_evening)
  
  results_evening <- data.frame(
    coef_name = names(betas_evening),
    beta = betas_evening,
    se = ses_evening,
    ci_low = betas_evening - 1.96 * ses_evening,
    ci_high = betas_evening + 1.96 * ses_evening
  )
  
  betas_daytime <- coef(event_model_daytime)
  ses_daytime <- se(event_model_daytime)
  
  results_daytime <- data.frame(
    coef_name = names(betas_daytime),
    beta = betas_daytime,
    se = ses_daytime,
    ci_low = betas_daytime - 1.96 * ses_daytime,
    ci_high = betas_daytime + 1.96 * ses_daytime
  )
  
  results_morning = results_morning %>% mutate(year_rel = sapply(coef_name,function(x) {return(unlist(str_split(x,":"))[3])}) %>% as.numeric())
  results_morning[6,] = c("year_rel::-1:treated",0,0,0,0,-1)
  results_morning = results_morning %>% select(-coef_name) %>% mutate(across(everything(),as.numeric))
  
  results_evening = results_evening %>% mutate(year_rel = sapply(coef_name,function(x) {return(unlist(str_split(x,":"))[3])}) %>% as.numeric())
  results_evening[6,] = c("year_rel::-1:treated",0,0,0,0,-1)
  results_evening = results_evening %>% select(-coef_name) %>% mutate(across(everything(),as.numeric))
  
  results_daytime = results_daytime %>% mutate(year_rel = sapply(coef_name,function(x) {return(unlist(str_split(x,":"))[3])}) %>% as.numeric())
  results_daytime[6,] = c("year_rel::-1:treated",0,0,0,0,-1)
  results_daytime = results_daytime %>% select(-coef_name) %>% mutate(across(everything(),as.numeric))
  
  
  
  results_daytime$hour = "daytime"
  results_morning$hour = 'morning'
  results_evening$hour = "evening"
  
  results = rbind(results_daytime,results_evening,results_morning)
  
  ggplot(results, aes(x = year_rel, y = beta, color = hour, shape = hour)) +
    geom_hline(yintercept = 0, linetype = "dashed", color = "gray") +
    geom_point(size = 2, position = position_dodge2(0.5)) +
    geom_errorbar(aes(ymin = ci_low, ymax = ci_high), width = 0, position = position_dodge(0.5)) +
    geom_line(position = position_dodge2(0.5))+
    labs(
      x = "Year relative to reform",
      y = paste0("Estimated effect (β_k) on ",as.character(variable)),
      title = "Estimated effect by time periods"
    ) +
    theme_minimal() + 
    scale_color_manual(values = c("black","darkgrey","darkblue"))
}



plot_occupancy = compute_model_plot("log_occupancy")
plot_flow = compute_model_plot("log_flow")
plot_speed = compute_model_plot("log_speed")
plot_congestion = compute_model_plot('congestion')


```

```{r}


library(dplyr)
library(zoo)      # pour rollmean
library(ggplot2)

Final_na <- Final %>%
  mutate(log_occupancy = na_if(log_occupancy, -Inf))

df_monthly <- Final_na %>%
  mutate(month = floor_date(date, "month")) %>%   # transforme la date en AAAA-MM-01
  group_by(iu_ac, month, treated) %>%             # agrégation mensuelle par route
  summarise(
    mean_log_occ = mean(log_occupancy, na.rm = TRUE),
    .groups = "drop"
  )

df_roll <- df_monthly %>%
  arrange(iu_ac, month) %>%
  group_by(iu_ac) %>%
  mutate(
    occ_roll12 = zoo::rollmean(
      mean_log_occ, 
      k = 12,
      align = "right",
      fill = NA,
      na.rm = TRUE
    )
  ) %>%
  ungroup()

df_roll_bis = df_roll %>% 
  group_by(treated, month) %>% 
  summarise(occ_roll12 = mean(occ_roll12,na.rm=TRUE))


ggplot(df_roll_bis, aes(x = month, y = occ_roll12, color = factor(treated))) +
  geom_line(size = 1.1) +
  theme_minimal(base_size = 14) +
  labs(
    title = "Common trends — 12-month rolling mean",
    x = "Month",
    y = "Rolling mean of log occupancy"
  )



```

